### 2014-07-03 Report/queue_minimizing_intro.tex
Update at home

% To begin with, we discuss the existing solution and its extension to solve the problem of resource allocation over multiple sub-channels. Then, we provide two approaches to solve the \ac{JSFRA} precoder design problem to minimize the total number of backlogged packets. At the end, we discuss a low complex solution for the \ac{JSFRA} problem by limiting the precoder design over each sub-channel at a given instant.

In general, the precoder design for the \ac{MIMO} \ac{OFDM} problem is highly difficult due to the combinatorial and the non-convex nature of the problem. In addition to that, the objective of minimizing the number of the queued packets over the spatial and the sub-channel dimensions adds further complexity to the existing problem. Since the scheduling of users in each sub-channel can be made by allocating zero transmit power over certain sub-channels, the solutions provided in the paper performs both precoder design and the scheduling of users in a joint manner.