### 2014-08-06 Report/simulation_intro.tex
Updated KKT MSE and figures

The simulations carried out in this work considered the path loss varying uniformly across all users in the system with the channels drawn from the i.i.d samples. The queues are generated based on the poisson process with the average values specified in each section presented.