### 2014-08-21 Report/simulation_intro.tex

The simulations carried out in this work considered the path loss varying uniformly across all users in the system with the channels drawn from the \textit{i.i.d} samples. The queues are generated based on the Poisson process with the average values specified in each section presented.