### 2014-10-02 Report/queue_minimizing_intro.tex
markku_modifications

% To begin with, we discuss the existing solution and its extension to solve the problem of resource allocation over multiple sub-channels. Then, we provide two approaches to solve the \ac{JSFRA} precoder design problem to minimize the total number of backlogged packets. At the end, we discuss a low complex solution for the \ac{JSFRA} problem by limiting the precoder design over each sub-channel at a given instant.

In general, the precoder design for the \ac{MIMO} \ac{OFDM} problem is highly difficult due to the combinatorial and the nonconvex nature of the problem. In addition, the objective of minimizing the number of the queued packets over the spatial and the sub-channel dimensions adds further complexity to the existing problem. Since the scheduling of users in each sub-channel can be made by allocating zero transmit power over certain sub-channels, our solutions perform joint precoder design and user scheduling. Before discussing the proposed solutions, we consider the existing algorithm to minimize the number of backlogged packets with additional constraints required by the problem.