### 2014-10-09 Report/joint_wsrm_formulation.tex
modified with markku input

The queue minimizing algorithms are discussed extensively in the networking literature to provide congestion-free routing between any two nodes in the network. One such algorithm  is the \emph{backpressure algorithm} \cite{tassiulas,georgiadis2006resource,neely2010stochastic}. It determines an optimal control policy in the form of rate or resource allocation for the nodes in the network by considering the differential backlogged packets between the source and the destination nodes. Even though the algorithm is primarily designed for the wired infrastructure, it can be extended to the wireless networks by designing the user rate variable \me{t_k} in accordance to the wireless network.

The \ac{Q-WSRM} formulation extends the \emph{backpressure algorithm} to the downlink \ac{MIMO}-\ac{OFDM} framework, in which the multiple \acp{BS} act as the source nodes and the user terminals as the receiver nodes. The control policy in the form of transmit precoders aims at minimizing the number of queued packets waiting in the \acp{BS}. In order to find the optimal strategy, we resort to the Lyapunov theory, which is predominantly used in the control theory to achieve system stability. Since at each time slot, the system is described by the channel conditions and the number of backlogged packets of each user, the Lyapunov function is used to provide a scalar measure, which grows large when the system moves toward the undesirable state. Following similar approach as in \cite{neely2010stochastic}, the scalar measure for the queue stability is given by
\begin{equation}
\mathrm{L}\sset{\mbf{Q}(i)} = \frac{1}{2} \, \sum_{k \in \mc{U}} Q_k^2(i),
\end{equation}
where \me{\mbf{Q}(i) = \sset{Q_1(i),Q_2(i),\dotsc,Q_K(i)}^\mathrm{T}} and \me{\tfrac{1}{2}} is used for the convenience. It provides a scalar measure of congestion present in the system \cite[Ch. 3]{neely2010stochastic}.

To minimize the total number of backlogged packets for an instant \me{i}, the optimal transmission rate of all users are obtained by minimizing the Lyapunov function drift expressed as  
\iftoggle{single_column}{
\begin{equation}\label{eqn-3.1}
\mathrm{L}\sset{\mbf{Q}(i+1)} - \mathrm{L}\sset{\mbf{Q}(i)} = \frac{1}{2} \Big [ \sum_{k \in \mc{U}} \, \Big ( \left [ Q_k(i) - t_k(i) \right ]^+ + \lambda_k(i) \Big )^2 - Q^2_k(i) \Big ].
\end{equation}}{
\begin{multline}\label{eqn-3.1}
\mathrm{L}\sset{\mbf{Q}(i+1)} - \mathrm{L}\sset{\mbf{Q}(i)} = \\ \frac{1}{2} \Big [ \sum_{k \in \mc{U}} \, \Big ( \left [ Q_k(i) - t_k(i) \right ]^+ + \lambda_k(i) \Big )^2 - Q^2_k(i) \Big ].
\end{multline}}
In order to eliminate the nonlinear operator \me{[x]^+}, we bound the expression in \eqref{eqn-3.1} as
\begin{equation}
\leq \sum_{k \in \mc{U}} \, \frac{\lambda^2_k(i) + t_k^2(i)}{2} + \sum_{k \in \mc{U}} Q_k(i) \set{ \lambda_k(i) - t_k(i) },
\label{drift-exp}
\end{equation}
by using the following inequality 
\begin{equation}
\sset{\max(Q-t,0) + \lambda)}^2 \leq Q^2 + t^2 + \lambda^2 + 2 Q (\lambda - t).
\end{equation}
The total number of backlogged packets at any given instant \me{i} is reduced by minimizing the conditional expectation of the Lyapunov drift expression \eqref{drift-exp} given the current number of queued packets \me{\mbfa{Q}(i)} waiting in the system. The expectation is taken over all possible arrival and transmission rates of the users to obtain the optimal rate allocation strategy.

Now, the conditional Lyapunov drift, denoted by \me{\Delta{(\mbfa{Q}(i))}}, is given by the infimum over the transmission rate as
\iftoggle{single_column}{
\begin{IEEEeqnarray}{rCl} \label{backpressure-approx}
	\underset{\mbf{t}}{\text{inf}} &\quad& \Delta(\mbf{Q}(i)) \triangleq \mathbb{E}_{\mbfa{\lambda},\mbf{t}} \set{\mathrm{L}\sset{\mbf{Q}(i+1)} - \mathrm{L}\sset{\mbf{Q}(i)} \vert \mbf{Q}(i)} \IEEEyessubnumber \\
	&\leq& \underbrace{\mathbb{E}_{\mbfa{\lambda},\mbf{t}} \set {\sum_{k \in \mc{U}} \, \frac{\lambda^2_k(i) + t_k^2(i)}{2} \vert \mbf{Q}(i)}}_{\le B} + \sum_{k \in \mc{U}} Q_k(i) A_k(i) - \mathbb{E}_{\mbfa{\lambda},\mbf{t}}\set{\sum_{k \in \mc{U}} Q_k(i) t_k(i)  \vert \mbf{Q}(i)}, \eqsub \label{backpressure-approx-1b}
\end{IEEEeqnarray}}{
\begin{IEEEeqnarray}{CL} \label{backpressure-approx}
	\underset{\mbf{t}}{\text{inf}} \quad & \mathbb{E}_{\mbfa{\lambda},\mbf{t}} \set{\mathrm{L}\sset{\mbf{Q}(i+1)} - \mathrm{L}\sset{\mbf{Q}(i)} \vert \mbf{Q}(i)} \IEEEyessubnumber \\
	\leq & \underbrace{\mathbb{E}_{\mbfa{\lambda},\mbf{t}} \Big \lbrace \sum_{k \in \mc{U}} \frac{\lambda^2_k(i) + t_k^2(i)}{2} \vert \mbf{Q}(i) \Big \rbrace }_{\le B} + \sum_{k \in \mc{U}} Q_k(i) A_k(i) \nonumber \\
	& \qquad \qquad {} - \mathbb{E}_{\mbfa{\lambda},\mbf{t}}\Big \lbrace \sum_{k \in \mc{U}} Q_k(i) t_k(i)  \vert \mbf{Q}(i) \Big \rbrace, \eqsub \label{backpressure-approx-1b}
\end{IEEEeqnarray}}
where the subscripts \me{\mbf{t}} and \me{\mbfa{\lambda}} represents the vector formed by stacking the transmission and the arrival rate of all users in the system. Since the transmission and the arrival rates are bounded, the second order moments in the first term of \eqref{backpressure-approx-1b} can be bounded by a constant \me{B} without affecting the optimal solution of the problem \cite{neely2010stochastic}. The second term in \eqref{backpressure-approx-1b} follows from the Poisson arrival process.

The expression in \eqref{backpressure-approx} looks similar to the \ac{WSRM} formulation if the weights in the \ac{WSRM} problem are replaced by the number of backlogged packets corresponding to the users. The above discussed approach is extended for the wireless networks in \cite{weeraddana2011resource}, where the queue weighted sum rate maximization is considered as the objective function to determine the transmit precoders. Since the expectation can be minimized by minimizing the function inside, the \ac{Q-WSRM} formulation is given by 
\begin{IEEEeqnarray}{LCl} \label{q_gen_sum}
	\underset{\substack{\mvec{M}{k,n}, \mvec{W}{k,n}}}{\text{maximize}} &\quad& \sum_{k \in \mc{U}} \, Q_k \left ( \sum_{n = 1}^N \sum_{l=1}^L \log_2(1 + \gamma_{l,k,n}) \right ) \IEEEyessubnumber \label{eqn-3.1.1} \\
	\text{subject to.} & \quad & \sum_{n = 1}^N \sum_{k \in \mathcal{U}_b} \text{tr} \, (\mvec{M}{k,n} \mvec{M}{k,n}^\herm) \leq P_{{\max}}, \fall b. \IEEEyessubnumber \label{eqn-3.1.3}
\end{IEEEeqnarray}

In order to avoid the excessive  allocation of the resources, we include an additional rate constraint \me{t_k \leq Q_k} to address \me{[x]^+} operation in \eqref{eqn-2a}. The rate constrained version of the \ac{Q-WSRM}, denoted by \ac{Q-WSRME} problem for a cellular system, is given by \label{q_gen_sum-1} with the additional constraint
\begin{IEEEeqnarray}{rCl} \label{eqn-3.1.4}
\sum_{n=1}^N \sum_{l = 1}^L \log_2(1 + \gamma_{l,k,n}) \leq Q_k, \fall k \in \mc{U},
\end{IEEEeqnarray}
where the precoders are associated with \me{\gamma_{l,k,n}} defined in \eqref{eq:SINR}. By using the number of queued packets as the weights, the resources can be allocated to the user with the more backlogged packets, which essentially does the allocation in a greedy manner.

As a special case of the problem defined in \eqref{q_gen_sum}, we can formulate the sum rate maximization problem by setting the weights in \eqref{eqn-3.1.1} as unity, leading to the problem as in \eqref{q_gen_sum} with \me{Q_k = 1, \forall k \in \mc{U}}. This approach provides a greedy queue minimizing allocation as compared to \ac{Q-WSRME}, since the resource allocation is driven by the channel conditions in comparison to the number of queued packets as in \ac{Q-WSRME}. Note that in both formulations, the resources allocated to the users are limited by the number of backlogged packets with an explicit maximum rate constraint defined by \eqref{eqn-3.1.4}.