### 2014-10-09 Report/simulation_intro.tex

The simulations carried out in this work consider the path loss varying uniformly across all users in the system with the channels drawn from the \textit{i.i.d.} samples. The queues are generated based on the Poisson process with the average values specified in each section presented.