### 2014-10-23 Report/conclusions.tex
final_submission_text

In this paper, we addressed the problem of allocating downlink space-frequency resources to the users in a multi-cell \ac{MIMO} \ac{IBC} system using \ac{OFDM} transmission. The resource allocation is considered as a joint space-frequency precoder design problem since the allocation of a resource to a user is obtained by a non-zero precoding vector. We proposed the \ac{JSFRA} scheme by adopting the \ac{SCA} technique to model the nonconvex constraint by a sequence of convex subsets to design the precoders for minimizing the total number of queued packets in the system. In addition to that, an alternative method using \ac{MSE} relaxation is also proposed using the same objective by fixing the receivers based on the \ac{MMSE} criterion. Then we proposed distributed precoder designs for the \ac{JSFRA} problem using primal and \ac{ADMM} methods. At last, we proposed a practical iterative algorithm to evaluate the precoders in a decentralized manner by solving the \ac{KKT} conditions for the \ac{MSE} reformulated \ac{JSFRA} scheme. The proposed iterative algorithm requires few iterations and limited signaling exchange between the coordinating \acp{BS} to obtain the precoders to minimize the total number of backlogged packets. Numerical results are used to compare the performance of the proposed algorithms with the existing solutions.