### 2014-10-24 Report/conclusions.tex
final version

In this paper, we addressed the problem of allocating downlink space-frequency resources to the users in a multi-cell \ac{MIMO} \ac{IBC} system using \ac{OFDM} transmission. The resource allocation is considered as a joint space-frequency precoder design problem since the allocation of a resource to a user is obtained by a non-zero precoding vector. We proposed the \ac{JSFRA} scheme by employing the \ac{SCA} technique to relax the nonconvex constraint by a sequence of convex subsets for designing the precoders to minimize the total number of user queued packets. Additionally, an alternative \ac{MSE} relaxation approach is also proposed by using \ac{SCA} technique to address the nonconvex constraints for a fixed \ac{MMSE} receivers. We then introduced distributed precoder designs for the \ac{JSFRA} problem using primal and \ac{ADMM} methods. Finally, we proposed a practical iterative algorithm to obtain the precoders in a decentralized manner by solving the \ac{KKT} conditions of the \ac{MSE} reformulated \ac{JSFRA} method. The proposed iterative algorithm requires few iterations and limited signaling exchange between the coordinating \acp{BS} to obtain the efficient precoders for a given number of iterations. Numerical results are used to compare the performance of the proposed algorithms with the existing solutions.