### 2014-10-24 Report/queue_minimizing_intro.tex
Submitted version

% To begin with, we discuss the existing solution and its extension to solve the problem of resource allocation over multiple sub-channels. Then, we provide two approaches to solve the \ac{JSFRA} precoder design problem to minimize the total number of backlogged packets. At the end, we discuss a low complex solution for the \ac{JSFRA} problem by limiting the precoder design over each sub-channel at a given instant.

In general, the precoder design for the \ac{MIMO} \ac{OFDM} problem is difficult due to the combinatorial and the nonconvex nature of the problem. In addition, the objective of minimizing the number of the queued packets over space-frequency dimensions adds further complexity to the existing problem. Since the scheduling of users in each space-frequency resource can be controlled by the transmit precoders, our solutions perform precoder design and user scheduling jointly. Before discussing the proposed solutions, we consider the existing algorithm to minimize the number of backlogged packets with additional constraints required by the problem.