### 2015-01-22 Report/appendix_tightness.tex
convergence_proof

In order for the constraints \eqref{eqn-6.2} and \eqref{eqn-6.3} to be active, there should be at least one user per \ac{BS} with large enough backlogged packets that cannot be satisfied with the given power budget. To prove this, let us consider a single \ac{BS} \me{N_T \times 1} system with \me{|U|_b} users each with \me{Q_k} backlogged packets. Let us consider the problem for a single \ac{BS} \me{b} as 
\begin{IEEEeqnarray}{rCl}
	\underset{\gamma_{k},\mvec{m}{k},\beta_{k}}{\text{minimize}} &\quad& \sum_{k \in \mc{U}_b} \big | Q_k - \log_2(1 + \gamma_k) \big |^q  \eqsub \\
	\text{subject to} &\quad& \gamma_{k} \leq \frac{ | \mvec{H}{k} \mvec{m}{k}|^2}{\beta_{k}} \IEEEyessubnumber \label{ae-1} \\
	&\quad& \beta_{k} \geq N_0 + \sum_{\mathclap{i \neq k}} |\mvec{H}{k} \mvec{m}{i} |^2 \IEEEyessubnumber \label{ae-2} \\
	&\quad& \sum_{k \in \mc{U}_b} \text{tr} \, (\mvec{m}{k} \mvec{m}{k}^\herm) \leq P_{{\max}}. \IEEEyessubnumber \label{ae-3}
\end{IEEEeqnarray}
Let us assume the system has \me{N_T > |\mc{U}_b|} and the precoders are determined for all users at the optimal point for which the objective is minimum and is nonzero. Let us consider two users, say \me{i} and \me{j} with precoders \me{\mvec{m}{i}} and \me{\mvec{m}{j}}. In order to show the tightness of the constraints \eqref{ae-1} and \eqref{ae-2}, let us assume by contradiction that \eqref{ae-1} and \eqref{ae-2} are not tight for a user \me{i} only. Let us also consider that \me{\gamma_k} for all users \me{k \in \mc{U}_b \backslash j} are identified to satisfy the backlogged packets.

Now, in order to minimize the objective of the \me{\ith{i}} user, \me{\gamma_i} should be \me{2^{Q_i} - 1} to have the minimum objective. Since the constraints \eqref{ae-1} and \eqref{ae-2} are not tight for the \me{\ith{i}} user, the actual \ac{SINR} on the r.h.s of \eqref{ae-1} is greater than \me{\gamma_i}. It is attributed to the excess power in the transmit precoder of user \me{i} as \me{\mvec{m}{i} = \bar{\mbf{m}}_{i} + \mbf{V}_i \mbf{e}}, where \me{\mvec{V}{i} = \mc{N}([\mbf{H}^{\mathrm{T}}_0, \dotsc, \mbf{H}^{\mathrm{T}}_{i-1},\mbf{H}^{\mathrm{T}}_{i+1}, \dotsc,\mbf{H}^{\mathrm{T}}_{|\mc{U}_b| - 1}]^{\mathrm{T}})} denotes the null space corresponding to the other user channel vectors and \me{\mbf{e}} be any random vector with power \me{\Delta P}. Note that the extra power has no impact on the interference constraint \eqref{ae-2} of other users.

Since the allocated rate of user \me{j} is less than the queued packets, we can find a precoder \me{\mvec{m}{j}} with the additional power of \me{\Delta P} as \me{\mvec{m}{j}^\prime = \bar{\mbf{m}}_{j} + \mvec{V}{j} \mbf{e}^\prime}. Note that the new precoder has no impact on the interference constraint of other users. The newly found precoder \me{\mvec{m}{j}^\prime} minimize the queue of user \me{j} without affecting other user rates, there by reduces the objective further. It is in contradiction to our original assumption that the power is utilized for the user \me{i} making the constraints loose. It can be seen that the constraints are tight as long as there is one user with unserviced backlogged packets at each \ac{BS} with the given power budget. Since the objective is not to minimize the power, when the power budget is more than sufficient to service the backlogged packets of all users, the \ac{JSFRA} scheme is not guaranteed to find the minimum power precoders to minimize the current backlogged packets in the system.