### 2015-02-26 Report/ieee_journal.tex
included antti comments

\documentclass[10pt,journal,twocolumn,letterpaper]{./../Styles/IEEEtran}

\usepackage[section]{placeins}
\usepackage{amsmath,amssymb}
\usepackage{acronym,etoolbox}
\usepackage{graphicx,epstopdf}
\usepackage[ruled]{algorithm2e}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage[subrefformat=parens,labelformat=parens,labelseparator=period,caption=false,font=footnotesize]{subfig}
\usepackage{subeqnarray,multirow,cite,array,setspace,fixltx2e,lineno,verbatim,mathtools}

\graphicspath{{./../Figures/}{./../Figures/Linux/}{./../Figures/Review/}}
\DeclareGraphicsExtensions{.eps}

\ifCLASSINFOpdf
\epstopdfsetup{update,prepend,prefersuffix=false,suffix=}
\DeclareGraphicsRule{.eps}{pdf}{.pdf}{`epstopdf #1}
\pdfcompresslevel=9
\fi

\input{shortcuts}
\input{acronyms}

\newtoggle{includereview}
\newtoggle{enable_dist_fig}
\newtoggle{single_column}
\togglefalse{single_column}
\togglefalse{enable_dist_fig}
\togglefalse{includereview}

\iftoggle{single_column}{
	\doublespacing
}

\title{Traffic Aware Resource Allocation Schemes for Multi-Cell \acs{MIMO}-\acs{OFDM} Systems}
\author{
	\IEEEauthorblockN{Ganesh Venkatraman \IEEEmembership{Student Member,~IEEE}\thanks{This work has been supported by the Finnish Funding Agency for Technology and Innovation (Tekes), Nokia Solutions Networks, Xilinx Ireland, Academy of Finland. Part of this work is presented in ICASSP 2014.}, Antti T\"{o}lli \IEEEmembership{Member,~IEEE}, Le-Nam Tran \IEEEmembership{Member,~IEEE}, and Markku Juntti \IEEEmembership{Senior Member,~IEEE}}\thanks{
	The authors are with the Centre for Wireless Communications (CWC), Department of Communications Engineering (DCE), University of Oulu, Oulu, FI-90014, (e-mail: \{ganesh.venkatraman, antti.tolli, le.nam.tran, markku.juntti\}@ee.oulu.fi).}}

\begin{document}

\iftoggle{includereview}{
\newpage
\onecolumn
\doublespacing
\section{Response to reviewer comments}
\input{review_comments}
\twocolumn
\newpage
\singlespacing
\setcounter{page}{1}
}

\maketitle

\begin{abstract}
\input{abstract}
\end{abstract}

\begin{IEEEkeywords}
Convex approximations, \ac{MIMO}-\acs{IBC}, \ac{MIMO}-\ac{OFDM}, Precoder design, \acs{SCA}, \ac{WSRM}.
\end{IEEEkeywords}

\acresetall

\section{Introduction} \label{sec-1}
\input{introduction}

\section{System Model and Problem Formulation} \label{sec-2-3.2}
\subsection{System Model} \label{sec-2}
\input{system_model}
\subsection{Problem Formulation} \label{sec-3.2}
\input{problem_formulation_intro}

\section{Proposed Queue Minimizing Precoder Designs} \label{sec-3}
\input{queue_minimizing_intro}

\subsection{Queue Weighted Sum Rate Maximization (\acs{Q-WSRM}) Formulation} \label{sec-3.1}
\input{joint_wsrm_formulation}

\subsection{\acs{JSFRA} Scheme via \ac{SCA} approach} \label{sec-3.2.1}
\input{problem_formulation}
\input{wsrm_formulation}

\subsection{\ac{JSFRA} Scheme via \ac{MSE} Reformulation} \label{sec-3.3}
\input{mse_formulation}

\subsection{Reduced Complexity Spatial Resource Allocation (\acs{SRA})} \label{sec-3.4}
\review{\input{sub_channel_wsrm}}

\section{Distributed Solutions} \label{sec-4}

\review{\input{decomposition_solutions}}

%\subsection{Decomposition based Approaches} \label{sec-4a}

\subsection{Primal Decomposition} \label{sec-4.1}
\input{primal_approach}
%\input{primal_algorithm}

\subsection{Alternating Directions Method of Multipliers (ADMM)} \label{sec-4.2}
\input{dual_approach}
%\input{dual_algorithm}

\subsection{Convergence Analysis for Distributed Algorithms} \label{sec-dist-conv}
\review{\input{convergence_distributed}}

\subsection{Decomposition via \ac{KKT} Conditions for \ac{MSE} Formulation} \label{sec-4.3}
\input{kkt_mse_approach}

\section{Simulation Results} \label{sec-5}
\input{simulation_intro}

\subsection{Centralized Solutions}
\input{simulation_results}

\subsection{Distributed Solutions}
\input{simulation_distributed}

\subsection{Average Backlogged Packets Over Slots} \label{time-correlated}
\review{\input{time_correlated}}

%acresetall
\section{Conclusions} \label{sec-6}
\input{conclusions}

\appendices

%\section{\acl{PD} approach} \label{a-2}
%\input{primal_approach_appendix}

\section{Tightness of \ac{SINR} relaxation} \label{a-3}
\review{\input{appendix_tightness}}

\section{Convergence Proof for Centralized Algorithm} \label{sec-3.5}
\review{\input{convergence_centralized}}

\section{\ac{KKT} conditions for \ac{MSE} approach} \label{a-1}
\input{kkt_appendix}

\bibliographystyle{./../Styles/IEEEtran}
\bibliography{./../Library/kirja_survey}

\end{document}