### 2015-03-02 Report/ieee_journal.tex
Merge remote-tracking branch 'origin/revision_tsp_1' into revision_tsp_1
Conflicts:
	Report/ieee_journal.dvi
	Report/ieee_journal.log
	Report/ieee_journal.pdf
	Report/ieee_journal.synctex.gz
	Report/ieee_journal.tex

\documentclass[10pt,journal,twocolumn,letterpaper]{./../Styles/IEEEtran}

\usepackage[section]{placeins}
\usepackage{amsmath,amssymb}
\usepackage{acronym,etoolbox}
\usepackage{graphicx,epstopdf}
\usepackage[ruled]{algorithm2e}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage[subrefformat=parens,labelformat=parens,labelseparator=period,caption=false,font=footnotesize]{subfig}
\usepackage{subeqnarray,multirow,cite,array,setspace,fixltx2e,lineno,verbatim,mathtools}

\graphicspath{{./../Figures/}{./../Figures/Linux/}{./../Figures/Review/}}
\DeclareGraphicsExtensions{.eps}

\ifCLASSINFOpdf
\epstopdfsetup{update,prepend,prefersuffix=false,suffix=}
\DeclareGraphicsRule{.eps}{pdf}{.pdf}{`epstopdf #1}
\pdfcompresslevel=9
\fi

\input{shortcuts}
\input{acronyms}

\newtoggle{includereview}
\newtoggle{enable_dist_fig}
\newtoggle{single_column}
\togglefalse{single_column}
\togglefalse{enable_dist_fig}
\togglefalse{includereview}

\iftoggle{single_column}{
	\doublespacing
}

\title{Traffic Aware Resource Allocation Schemes for Multi-Cell \acs{MIMO}-\acs{OFDM} Systems}
\author{
	\IEEEauthorblockN{Ganesh Venkatraman$^\dagger$ \IEEEmembership{Student Member,~IEEE}\thanks{This work has been supported by the Finnish Funding Agency for Technology and Innovation (Tekes), Nokia Solutions Networks, Xilinx Ireland, Academy of Finland. Part of this work is presented in ICASSP 2014.}, Antti T\"{o}lli$^\dagger$ \IEEEmembership{Senior Member,~IEEE}, Le-Nam Tran$^\ddagger$ \IEEEmembership{Member,~IEEE}, and Markku Juntti$^\dagger$ \IEEEmembership{Senior Member,~IEEE}}\thanks{
	$^\dagger$The authors are with the Centre for Wireless Communications (CWC), Department of Communications Engineering (DCE), University of Oulu, Oulu, FI-90014, (e-mail: \{ganesh.venkatraman, antti.tolli, markku.juntti\}@ee.oulu.fi)}\thanks{$^\ddagger$The author is with the Department of Electronic Engineering, Maynooth University, Maynooth, Co. Kildare, Ireland, (e-mail:ltran@eeng.nuim.ie}}

\begin{document}

\iftoggle{includereview}{
\newpage
\onecolumn
\doublespacing
\section{Response to reviewer comments}
\input{review_comments}
\twocolumn
\newpage
\singlespacing
\setcounter{page}{1}
}

\maketitle

\begin{abstract}
\input{abstract}
\end{abstract}

\begin{IEEEkeywords}
Convex approximations, \ac{MIMO}-\acs{IBC}, \ac{MIMO}-\ac{OFDM}, Precoder design, \acs{SCA}, \ac{WSRM}.
\end{IEEEkeywords}

\acresetall

\section{Introduction} \label{sec-1}
\input{introduction}

\section{System Model and Problem Formulation} \label{sec-2-3.2}
\subsection{System Model} \label{sec-2}
\input{system_model}
\subsection{Problem Formulation} \label{sec-3.2}
\input{problem_formulation_intro}

\section{Proposed Queue Minimizing Precoder Designs} \label{sec-3}
\input{queue_minimizing_intro}

\subsection{Queue Weighted Sum Rate Maximization (\acs{Q-WSRM}) Formulation} \label{sec-3.1}
\input{joint_wsrm_formulation}

\subsection{\acs{JSFRA} Scheme via \ac{SCA} approach} \label{sec-3.2.1}
\input{problem_formulation}
\input{wsrm_formulation}

\subsection{\ac{JSFRA} Scheme via \ac{MSE} Reformulation} \label{sec-3.3}
\input{mse_formulation}

\subsection{Reduced Complexity Spatial Resource Allocation (\acs{SRA})} \label{sec-3.4}
\review{\input{sub_channel_wsrm}}

\section{Distributed Solutions} \label{sec-4}

\review{\input{decomposition_solutions}}

%\subsection{Decomposition based Approaches} \label{sec-4a}

\subsection{Primal Decomposition} \label{sec-4.1}
\input{primal_approach}
%\input{primal_algorithm}

\subsection{Alternating Directions Method of Multipliers (ADMM)} \label{sec-4.2}
\input{dual_approach}
%\input{dual_algorithm}

\subsection{Convergence Analysis for Distributed Algorithms} \label{sec-dist-conv}
\review{\input{convergence_distributed}}

\subsection{Decomposition via \ac{KKT} Conditions for \ac{MSE} Formulation} \label{sec-4.3}
\input{kkt_mse_approach}

\section{Simulation Results} \label{sec-5}
\input{simulation_intro}

\subsection{Centralized Solutions}
\input{simulation_results}

\subsection{Distributed Solutions}
\input{simulation_distributed}

\subsection{Queue Deviations over Multiple Transmission Slots} \label{time-correlated}
\review{\input{time_correlated}}

%acresetall
\section{Conclusions} \label{sec-6}
\input{conclusions}

\appendices

%\section{\acl{PD} approach} \label{a-2}
%\input{primal_approach_appendix}

\section{Tightness of \ac{SINR} relaxation} \label{a-3}
\review{\input{appendix_tightness}}

\section{Convergence Proof for Centralized Algorithm} \label{sec-3.5}
\review{\input{convergence_centralized}}

\section{\ac{KKT} conditions for \ac{MSE} approach} \label{a-1}
\input{kkt_appendix}

\bibliographystyle{./../Styles/IEEEtran}
\bibliography{./../Library/kirja_survey}

\end{document}