### 2015-03-03 Response/reviewer_response.tex

\documentclass[10pt,letterpaper,onecolumn]{IEEEtran}

\usepackage[left=1in,right=1in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{acronym,etoolbox}
\usepackage{subeqnarray,multirow,cite,array,setspace,fixltx2e,lineno,verbatim,mathtools}

\newcommand{\cmnt}[1]{\item[\textbf{Q}-(#1)]}
\newcommand{\cmnts}[1]{\item[(#1)]}
\newcommand{\resp}{\item[\textbf{Resp} :]}

\renewcommand \thesection{\Roman{section}}
\renewcommand \thesubsection{\arabic{section}.\arabic{subsection}}

\input{./../Report/shortcuts}
\input{./../Report/acronyms}

\begin{document}

\doublespacing

\title{Response for Reviewer Comments}

\maketitle

We would like to thank the associate editor and the reviewers for the valuable comments provided for the manuscript. We have written the response for the reviewer comments in line in this report. The changes are addressed in the revised manuscript as well by using blue color italic fonts to discriminate the revision from the actual text in the manuscript.

\newpage
\section*{Response for Reviewer - \me{1}}
\input{response-1}

\newpage
\section*{Response for Reviewer - \me{2}}
\input{response-2}

\newpage
\section*{Response for Reviewer - \me{3}}
\input{response-3}

\newpage
\section*{Response for Reviewer - \me{4}}
\input{response-4}

\end{document}