### 2015-03-19 Report/conclusions.tex

In this paper, we addressed the problem of allocating downlink space-frequency resources to the users in a multi-cell \ac{MIMO} \ac{IBC} system using \ac{OFDM}. The resource allocation is considered as a joint space-frequency precoder design problem since the allocation of a resource to a user is obtained by a non-zero precoding vector. We proposed the \ac{JSFRA} scheme by relaxing the nonconvex \ac{DC} constraint by a sequence of convex subsets using the \ac{SCA} for designing the precoders to minimize the total number of user queued packets. Additionally, an alternative \ac{MSE} relaxation approach is also proposed by using the \ac{SCA} to address the nonconvex \ac{DC} constraints for a fixed \ac{MMSE} receivers. We also proposed distributed precoder designs for the \ac{JSFRA} problem using primal and \ac{ADMM} methods. Finally, we proposed a practical iterative algorithm to obtain the precoders in a decentralized manner by solving the \ac{KKT} conditions of the \ac{MSE} reformulated \ac{JSFRA} method.  The proposed iterative algorithm requires few iterations and limited signaling exchange between the coordinating \acp{BS} to obtain the efficient precoders for a given number of iterations. Numerical results are used to compare the performances of the proposed algorithms with the existing solutions. The distributed design for the time correlated fading will be considered in future.