### 2015-03-19 Report/introduction.tex

%The last mile wireless connectivity poses significant bottleneck in the overall data traffic for the interconnected networks. The main challenges in the wireless networks are due to the scarcity of the available resources either in terms of power or spectrum usage and the complexity of the receiver algorithms which impacts the mobile battery consumption. In order to overcome the receiver complexity, \ac{OFDM} is introduced for the wide band transmissions. To improve data rate, multiple antennas are installed at \acp{BS} and/or at user terminals to avail additional freedom in the form of spatial dimension. The inclusion of \ac{MIMO} technique in wireless networks provides higher data rate or lower outage for the same transmission power and bandwidth.

\acused{MU}

In a network with multiple \acp{BS} serving multiple users, the main driving factor for the transmission are the packets waiting at each \ac{BS} corresponding to the different users present in the network. We consider the problem of transmit precoder design over the space-frequency resources provided by the \ac{MIMO} \ac{OFDM} framework in the downlink \ac{IBC} to minimize the number of queued packets at each \ac{BS}. Since the available resources are shared by multiple users associated with different \acp{BS}, it can be viewed as a resource allocation problem.

In general, the resource allocation problems can be formulated by assigning a binary variable for each user to indicate the presence or the absence in a particular resource \cite{admission_control}. \review{In contrast, the linear transmit precoders, which are complex vectors, are implicitly used as decision variables, thereby avoiding the explicit modeling of binary decision variables. They are used to determine the transmission rates of users on a space-frequency resource if the precoder vector is non-zero. A zero transmit precoder indicates the absence of the user on a given resource. In this way, the soft decisions are used in the optimization problem and the hard decisions are made after the algorithm convergences}.

%The queue minimizing network optimization objective is used to design the beamformers across the coordinating \acp{BS}, since the transmissions are guided by the available backlogged packets. To achieve the best performance, we propose a joint resource allocation scheme over the space and frequency dimensions among the coordinating \acp{BS} to minimize the time that the packets stay in the queues prior to the transmission, and, hence, to avoid packet drops as an indirect objective.

The queue minimizing precoder designs are closely related to the \ac{WSRM} problem with additional rate constraints determined by the number of backlogged packets for each user in the system. The topics on \ac{MIMO} \ac{IBC} precoder design have been studied extensively with different performance criteria in the literature. Due to the nonconvex nature of the \ac{MIMO} \ac{IBC} precoder design problems, the \ac{SCA} approach has become a powerful tool to deal with these problems. For example, in \cite{sin_algorithm}, the nonconvex part of the objective has been linearized around an operating point in order to solve the \ac{WSRM} problem in an iterative manner. Similar approach of solving the \ac{WSRM} problem by using arithmetic-geometric inequality has been proposed in \cite{tran2012fast}.

The relation between the achievable sum rate and the \ac{MSE} of the received symbol by using fixed \ac{MMSE} receivers can be used to solve the \ac{WSRM} problem \cite{mse_duality}. In \cite{christensen2008weighted,wmmse_shi}, the \ac{WSRM} problem is reformulated via \ac{MSE}, casting the problem as a convex one for fixed linearization coefficients. In this way, the original problem is expressed in terms of the \ac{MSE} weight, precoders, and decoders. Then the problem is solved using an alternating optimization method, i.e., finding a subset of variables while the other variables are fixed. The \ac{MSE} reformulation for the \ac{WSRM} problem has also been studied in \cite{hong2012decomposition} by using the \ac{SCA} to solve the problem in an iterative manner. Additional the rate constraints based on the \ac{QoS} requirements were included in the \ac{WSRM} problem and solved via the \ac{MSE} reformulation in \cite{kaleva2013primal,kaleva2013decentralized}.

The problem of precoder design for the \ac{MIMO} \ac{IBC} system can be solved either by using a centralized controller or by using decentralized algorithms, where each \ac{BS} handles the corresponding subproblem independently with the limited information exchange with other \acp{BS} via back-haul. The distributed approaches are usually based on the primal, or dual decompositions or the \ac{ADMM}, which has been discussed in \cite{palomar2006tutorial,boyd2011distributed}. In the  primal decomposition, the so-called coupling interference variables are fixed for the subproblem at each \ac{BS} to find the optimal precoders. The fixed interference is then updated by using the subgradient method as discussed in \cite{pennanen2011decentralized}. The dual and the \ac{ADMM} approaches control the distributed subproblems by fixing the \emph{`interference price'} for each \ac{BS} as detailed in \cite{tolli2011decentralized}.

By adjusting the weights in the \ac{WSRM} objective properly, we can find an arbitrary rate-tuple in the rate region that maximizes the suitable objective measures. For example, if the weight of each user is set to be inversely proportional to its average data rate, the corresponding problem guarantees fairness on the average among the users. As an approximation, we may assign weights based on the current queue size of the users. Specifically, the queue states can be incorporated in the \ac{WSRM} objective $\sum_k w_k R_k$ by replacing the weight $w_k$ with the corresponding queue state $Q_k$ or its function, which is the outcome of minimizing the Lyapunov drift between the current and the future queue states \cite{tassiulas,neely2010stochastic}, where \me{R_k} denotes the achievable user rate. In the backpressure algorithm, the differential queues between the source and the receiver nodes are used as the weights scaling the transmission rate \cite{georgiadis2006resource}.

Earlier studies on the queue minimization problem were summarized in the survey papers \cite{berry2004cross,layering_as_opt}. In particular, the problem of power allocation to minimize the number of backlogged packets was considered in \cite{qps_cioffi} using geometric programming. Since the problem addressed in \cite{qps_cioffi} assumed single antenna transmitters and receivers, the queue minimizing problem reduces to the optimal power allocation problem. In the context of wireless networks, the backpressure algorithm mentioned above was extended in \cite{weeraddana2011resource} by formulating the corresponding user queues as the weights in the \ac{WSRM} problem. Recently, the precoder design for the video transmission over \ac{MIMO} system is considered in \cite{video_queues}. In this design, the \ac{MU}-\ac{MIMO} precoders are designed by the \ac{MSE} reformulation as in \cite{christensen2008weighted} with the higher layer performance objective such as playback interruptions and buffer overflow probabilities.

\subsubsection*{Main Contributions}
\review{In this paper, we design the precoders jointly across space-frequency resources by minimizing the total number of backlogged packets waiting at the \acp{BS} queues. Since the transmissions are guided by the backlogged packets, the proposed formulation limits the allocations beyond the number of backlogged packets without additional rate constraint. We adopt the \ac{SCA} to solve a sequence of convex subproblems using the first order Taylor approximation due to the nonconvexity imposed by the \ac{DC} constraints.} Initially, we propose centralized \ac{JSFRA} algorithms, which employs the \ac{SCA} for the nonconvex \ac{DC} constraint. The first method is by using the \ac{SINR} expression and the second one is by using the \ac{MSE} equivalence with the rate expression to solve for the optimal precoders. Then we propose distributed precoder designs based on the primal and the \ac{ADMM} methods. Finally, an iterative practical algorithm is proposed to decouple the precoder design across the coordinating \acp{BS} with limited information exchange by solving the \ac{KKT} conditions for the \ac{MSE} reformulation. \review{Note that the joint space-frequency channel matrix can be formed by stacking the channel of each sub-channel in a block-diagonal form for all users}.

The rest of the paper is organized as follows. In Section \ref{sec-2-3.2}, we introduce the system model and the problem formulation for the queue minimizing precoder design. The existing and the proposed centralized precoder designs are presented in Section \ref{sec-3}. The distributed solutions are provided in Section \ref{sec-4} followed by the simulation results in Section \ref{sec-5}. Conclusions are drawn in Section \ref{sec-6}.