### 2015-04-07 Response/additions.tex

\begin{enumerate}
\item We have provided additional information to improve the continuity in the algorithm formulation from the SINR expression as suggested by the reviewers.
\item We have included the discussion for the MSE based reformulation for different norms in the objective function.
\item We have rewritten the reduced complexity spatial resource allocation (SRA) in Section III-D for better readability.
\item We have shortened the ADMM approach as suggested by the reviewers in Section IV-B.
\item Rigorous convergence analysis is provided for the centralized algorithm in Appendix A.
\item We have included Section V-C on the queue behavior over multiple time instants. We have added Fig. 4 to show the performance of the proposed scheme over the existing precoder design algorithms.
\item We have addressed the issues in the citations and also provided additional references to prove the convergence of the centralized algorithm as suggested by the reviewers.
\end{enumerate}

Due to the page limitation imposed on the resubmitted manuscript, we have included all the Appendices as a supplementary document in the manuscript central. In what follows, the comments are listed, each followed by the corresponding reply from the authors. Unless otherwise stated, all the numbered items (figures, equations, references, citations, etc) in this response letter refer to the revised manuscript.