### 2015-06-22 Response/Response 2/response-3.tex
updated

The authors have introduced changes in the manuscript that improved the paper's quality. Additionally, the authors have taken into account the reviewers' comments giving clarifications and modifying the content when required. More specifically, the following aspects have been treated:

\vspace{0.1in}
\review{We thank the reviewer for reading the manuscript and providing valuable comments. The comments are really helps to improve the manuscript.}

\begin{enumerate}
\cmnt{1} Convexity of problem (16). The paragraphs surrounding (16) allow a better understanding of the usage of the additional variables, i.e. gamma and beta, to remove the equality constraint in (2). For the reviewer remains however unclear, the procedure/criterion to determine the operating point for the parameter \eqn{\tilde{\beta}} required in (19) and used in the convex subproblems (20) and (21).

\resp \review{
	}

\cmnt{2} Proof of convergence. The proof of convergence introduced by the authors in the Appendix seems correct and enhances the content of the manuscript.

\resp \review{}

Additional Comments -

\begin{enumerate}

	\cmnts{a} - {The reviewer considers that closing statements regarding the applicability of the proposed schemes are missing. Since the results are quite similar (when not identical), which formulation is preferable between the centralized schemes? Which one for the distributed solutions?}

	\resp \review{}

	\cmnts{b} - {The last discussion in section IV-C could benefit from restructuring. The information on how to obtain a practical distributed precoder design and to avoid backhaul exchange is too condensed and difficult to understand.}

	\resp \review{}

	\cmnts{c} - {For the simulation results, why not to unify configurations when possible? Having to read a different configuration for each graph is cumbersome and no additional comparisons are possible between figures. E.g. PL uniformly distributed between [0,-6] dB in Fig. 2 and [0,-3] dB in Fig. 3.}

	\resp \review{}

	\cmnts{d} - {In Fig. 1, the description of the system model does not agree with the statement of N = 3 subchannels.}

	\resp \review{}

	\cmnts{e} - {In Fig. 4(b), the performance of Q-WSRME seems to be (in average) worse than Q-WSRM. However, that should not be the case, since Q-WSRME is taking into account the over allocation. Any reason for this?}

	\resp \review{}

	\cmnts{f} - {p 6, col 2, row 49: it should be \eqn{t_{l,k,n}} instead of \eqn{t_{l,n,k}}}

	\resp \review{}

	\cmnts{g} - {p 10, col 1, row 28: is it \eqn{\lambda} a dual variable?}

	\resp \review{}

	\cmnts{h} - {p 10, col 2, row 59: typo wHith}

	\resp \review{}

	\cmnts{i} - {In general, a grammar check is recommended, several mistakes with respect to singular and plural nouns have been observed, e.g. -p 6, col 2, row 56 "... for A fixed receiverS" is not correct.}

	\resp \review{}

\end{enumerate}

\end{enumerate}