### 2015-06-27 Response/Response 2/additions.tex

\begin{enumerate}
\item We have shortened the discussions on the existing \textit{backpressure algorithm} in Section III-A.
\item Strong convexity of the objective function is emphasized in Appendix A-C.
\item Strict Monotonicity of the algorithm is discussed separately in Appendix A-D with more clarity.
\item We have provided additional details on the distributed algorithm convergence as suggested by the reviewers
\item Section III-D is updated to include the guidelines involved in selecting the sub-channel order.
\item Short discussion on the choice of algorithm selection is provided 
\item Initialization of the \ac{SCA} operating point is also included in the revised manuscript
\end{enumerate}

In what follows, the comments are listed, each followed by the corresponding reply from the authors. Unless otherwise stated, all the numbered items (figures, equations, references, citations, etc) in this response letter refer to the revised manuscript.