### 2015-06-27 Response/Response 2/document.tex

\documentclass[10pt,letterpaper,onecolumn]{article}

\usepackage[left=1.0in,right=1.0in,top=1in,bottom=1in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage[dvipdfmx]{graphicx}
\usepackage{acronym,etoolbox,color,multicol,xcolor}
\usepackage{subeqnarray,multirow,cite,array,setspace,fixltx2e,lineno,verbatim,mathtools}

\newcommand{\cmnt}[1]{\item}
\newcommand{\cmnts}[1]{\item[(#1)]}
\newcommand{\resp}{\item[\textit{\underline{Reply}}:]}

\input{./../../Report/shortcuts}
\input{./../../Report/acronyms}

\DeclareGraphicsExtensions{{eps}}
\graphicspath{{Figures/}}
\renewcommand \thesection{\Roman{section}}
\renewcommand \thesubsection{\arabic{section}.\arabic{subsection}}

\begin{document}

\title{Reviewers Comments \& Authors Replies}

\date{}
\maketitle

\begin{tabular}{p{1.25in}p{4.25in}}
\textbf{Manuscript No.} & Paper T-SP-18051-2014, submitted to \emph{``IEEE Transactions on Signal Processing"} \\ \\
\textbf{Title} & ``Traffic Aware Resource Allocation Schemes for Multi-Cell MIMO-OFDM Systems" \\ \\
\textbf{Authors} & Ganesh Venkatraman, Antti T\"{o}lli, Markku Juntti, and Le-Nam Tran
\end{tabular}

\vspace{0.5in}
The authors would like to thank the associate editor and the reviewers for their valuable comments on the manuscript of the paper, which have been greatly helpful to improve the paper quality. Based on the comments, we have made several major revisions to the paper, following the suggestions of the reviewers. In what follows, the comments are listed, each followed immediately by the corresponding reply from the authors. The revisions in the revised manuscript are highlighted using blue color and the authors responses are also presented in blue color text. Following is the summary of the revision made on the manuscript in accordance with the reviewers comments.
\input{additions}

\newpage
\section*{Response to Reviewer - \me{1}'s Comments}
\input{response-1}

\newpage
\section*{Response to Reviewer - \me{2}'s Comments}
\input{response-2}

\newpage
\section*{Response to Reviewer - \me{3}'s Comments}
\input{response-3}

\newpage
\section*{Response to Reviewer - \me{4}'s Comments}
\input{response-4}

\end{document}