### 2015-07-16 Response/Response 2/response-1.tex
updated with nam response

Comments:
\review{The response to the reviewer's concerns are generally satisfying, except the convergence proof.}

\vspace{1eM}
\underline{\textit{Reply:}} We thank the reviewer for providing valuable and insightful comments.

\vspace{1eM}
For the resubmitted manuscript, the reviewer still has the following concerns

\begin{enumerate}
\cmnt{1} \review{Considering the length of the manuscript, it would be better to shorten some parts that are not new in this manuscript, e.g. III.A. More space can be left for convergence proof, which is very important. }

\resp After a careful check of the manuscript, we have removed couple of paragraphs from Section III.A and shortened the discussions on the simulation section to provide additional details for the convergence proof. We believe the removed parts don't affect the readability of the paper and the added text regarding convergence proof certainly improves the quality.

\cmnt{2} \review{In convergence proof (48), why does the 2nd inequality hold? In fact, to prove the feasibility of  \eqn{{m_{k+1}^{(i)}, w_{*}^{(i-1)};m_k^{(i)} }}, the part between 2nd and 3rd inequality is not necessary, \eqn{\leq 0} directly follows the 2nd inequality since the solution \eqn{m_{k+1}^{(i)}, \gamma_{k+1}^{(i)} } is the optimal solution, and therefore feasible.  }

\resp We thank the reviewer for the critical comment. It is not possible to define the inequality with the previous optimal point. Since it is not possible to comment on the inclusion of the previous constraint set in the current update (except the earlier optimal point), the inequality is not guaranteed. Based on the comment from the reviewer, we have removed the inequality specifying the previous operating point from (49) in the revised manuscript, since it cannot be guaranteed.

\cmnt{3} \review{The solutions SCA iterations \eqn{\mbf{m}^{(i)}_k} does not necessarily converge. In fact \eqn{\mbf{m}} has compact feasible region, and thus \eqn{\mbf{m}^{(i)}_k} has limit points for any specific \eqn{i}.  However \eqn{m_{*}^{(i)}} does not necessarily exist( the whole sequence \eqn{\mbf{m}^{(i)}_k} may be not convergent). Similar problem happens to \eqn{\mbf{w}^{(i)}_k}. }

\resp We understand the reviewer concern. Even though the objective of the \ac{SCA} iteration converges, it is not guaranteed that the iterates involved in the iterative algorithm, namely, \eqn{\mbf{m}^{(i)}_k} and \eqn{\mbf{w}^{(i)}_k} to converge. It is true that the iterates need not converge when the objective function is convex. We have modified the discussion by using the strong convexity of the objective function to impose the uniqueness of the iterates in each \ac{SCA} update as well. By regularizing the objective function with a strongly convex term like \eqn{\|\mbf{m} - \mbf{m}^{(i)}_k\|^2}, we can guarantee the uniqueness of the iterates upon the \ac{SCA} convergence, \textit{i.e.}, the limit point of the \ac{SCA} update is unique. We thank the reviewer for citing this issue on the sequence convergence and the uniqueness of the minimizer. We have updated the manuscript to include this information in the convergence proof in Appendix A-B around (46) and in Appendix A-C.

\cmnt{4} \review{Strict monotonicity with respect to the objective function \eqn{f} should be rigorously proved. Note that to guarantee the uniqueness of the beamformer iterates, (52) instead of the objective function is used.}

\resp We thank the reviewer for the pointing out the issue involving the strict monotonicity. As suggested by the reviewer, we have included the strict monotonicity of the objective sequence in Appendix A-C following (52).

Note that the objective sequence \eqn{\{f(\mbf{x}_k)\}} generated by an iterative \ac{SCA} algorithm is monotonic, \textit{i.e}, \eqn{f(\mbf{x}_k) \leq f(\mbf{x}_{k-1})} and holds with equality at the limit point of the iterative procedure. However, upon convergence, due to the convexity of the objective, we may have multiple limit points, say, \eqn{\mbf{y}_{\ast}} with the same objective as \eqn{f(\mbf{y}_{\ast}) = f(\mbf{x}_{\ast})}. Therefore, strict monotonicity cannot be guaranteed for the objective sequence. However, when the objective function is strongly convex, as discussed in Appendix A-B, the minimizer for the subproblem in each \ac{SCA} iteration \eqn{k} is unique as
\begin{equation} \label{eqn-aa}
f(\ma_{k}) - f(\ma_{k+1}) \geq  \nabla f(\ma_{k+1})^\tran (\ma_{k} - \ma_{k+1}) + c \|\ma_{k} - \ma_{k+1}\|^2
\end{equation}
where \eqn{\ma_{k+1}} is the optimal solution for the \eqn{\ith{k}} \ac{SCA} subproblem and \eqn{\ma_{k}} is the unique minimizer obtained in the previous \eqn{\ith{k-1}} \ac{SCA} step, which is a feasible point for the current problem. Since \eqn{\ma_{k+1}} is the solution in the \eqn{\ith{k}} \ac{SCA} problem, following conditions hold
\begin{subeqnarray}
\nabla f(\ma_{k+1})^\tran (\ma_{k} - \ma_{k+1}) &\geq& 0 \slabel{eqn-a} \\
f(\ma_{k}) - f(\ma_{k+1}) &\geq& c \|\ma_{k} - \ma_{k+1}\|^2 \slabel{eqn-b}
\end{subeqnarray}
where \eqref{eqn-a} is due to the lack of any descent direction in the feasible set and \eqref{eqn-b} follows from \eqref{eqn-aa} using \eqref{eqn-a}. As \eqn{k \rightarrow \infty, \|\ma_{k+1} - \ma_{k}\| \rightarrow 0}, and therefore the iterative algorithm converges to a unique minimizer, say, \eqn{\ma_{\ast}}, which is the limit point of the algorithm. Now by using the monotonicity of the \ac{SCA} updates and the uniqueness of the minimizer in each step, we can guarantee the strict monotonicity of the objective sequence generated by the iterative problem.

\cmnt{5} \review{Note that the conclusions [32, Thm 2] and [26, Thm 10] have lots of assumptions. To invoke these reference, explicit exposition should be provided to show that these conclusions can be applied to our problem. The same questions occur to the proof in Appendix B, where conclusions in [11] [36] and [37] are used. Too many details are omitted to make the proof convincing and clear. }

\resp We thank the reviewer for raising the concern. We have updated the manuscript to include the details regarding the stationary point discussion in Appendix A-E and the convergence proof analysis for the primal and the \ac{ADMM} algorithms in Appendix B. Additional detail includes the conditions required for showing the stationarity of the limit point and we have included additional details to show the distributed approaches convergence convincing.

\end{enumerate}