### 2015-07-20 Response/Response 2/additions.tex

\begin{enumerate}
\item We have shortened the discussions on the existing \textit{backpressure algorithm} in Section III-A.
\item Initialization of the \ac{SCA} operating point is included in the revised manuscript
\item Section III-D is updated to include the guidelines involved in selecting the sub-channel order.
\item Strong convexity of the objective function is emphasized in Appendix A-B.
\item Strict monotonicity of the algorithm is discussed in a detailed manner in Appendix A-C.
\item We have provided additional details on the distributed algorithm convergence as suggested by the reviewers
\end{enumerate}

In what follows, the comments are listed, each followed immediately by the corresponding reply from the authors. The reviewers questions in the revised manuscript are highlighted in blue color and the authors responses are presented in black. Unless otherwise stated, all the numbered items (figures, equations, references, citations, etc) in this response letter refer to the revised manuscript.