### 2015-11-05 Response/Response 3/additions.tex

\begin{enumerate}
	\item We have rewritten Appendix A-D to discuss the stationarity of limit points of the sequence of beamformer iterates generated by the proposed algorithm.
	\item We have updated the convergence of distributed algorithm to reflect the same.
	\item We included additional constraints on optimization variables to ensure compactness of feasible set in each subproblem
\end{enumerate}

In what follows, the comments are listed, each followed immediately by the corresponding reply from the authors. The reviewers questions in the revised manuscript are highlighted in blue color and the authors responses are presented in black. Unless otherwise stated, all the numbered items (figures, equations, references, citations, etc) in this response letter refer to the revised manuscript. All revisions in the manuscript are highlighted in blue color.