### 2015-11-19 Report/problem_formulation_intro.tex

In order to minimize the total number of backlogged packets, we consider minimizing the weighted \me{\ell_{q}}-norm of the queue deviation metric as
\begin{equation} \label{eqn-4.2}
v_k =  Q_k - t_k = Q_k - \sum_{n = 1}^N \sum_{l = 1}^{L} \log_2(1+\gamma_{l,k,n})
\end{equation}
where \me{\gamma_{l,k,n}} is given by \eqref{eq:SINR} and the optimization variables are transmit precoders \me{\mvec{m}{l,k,n}} and receive beamformers \me{\mvec{w}{l,k,n}}.

Explicitly, the objective of the problem considered is given as \me{\sum_{k \in \mc{U}} \, a_k | v_k |^q}. Thus, the formulation becomes
\begin{IEEEeqnarray}{rCl} \label{eqn-3} \eqsubn
\underset{\substack{\mvec{m}{l,k,n},\mvec{w}{l,k,n}}}{\text{minimize}} &\quad& \|  \tilde{\mbf{v}}  \|_q \label{eqn-3-1.a} \\
\text{subject to} & \quad&\sum_{n = 1}^N \sum_{k \in \mathcal{U}_b} \sum_{l=1}^L \trace \, (\mvec{m}{l,k,n} \mvec{m}{l,k,n}^\herm) \leq P_{{\max}} \fall b \eqspace \label{eqn-4.3} \end{IEEEeqnarray}
where \me{\tilde{v}_k \triangleq a_k^{1/{q}}v_k} is the element of vector \me{\tilde{\mbf{v}}}. Weighing factors \me{a_k} are used to alter the user priority based on the \ac{QoS} constraints such as packet delay requirements and packet waiting time, since they are proportional to the corresponding number of backlogged packets. The \ac{BS} specific power constraint for all sub-channels is considered in \eqref{eqn-4.3}.

For practical reasons, we impose a constraint on the maximum number of transmitted bits for the user \me{k}, since it is limited by the total number of backlogged packets available at the transmitter. As a result, the number of backlogged packets \me{v_k} for user \me{k} remaining in the system is given by
\begin{equation} \label{rate_constraint_a}
v_k =  Q_k - \sum_{n = 1}^N \sum_{l = 1}^{L} \log_2(1+\gamma_{l,k,n}) \geq 0.
\end{equation}
The above positivity constraint needs to be satisfied by \me{v_k} to avoid the excessive allocation of the resources.

Before proceeding further, we show that the constraint in \eqref{rate_constraint_a} is handled implicitly by the definition of \me{\ell_q} norm in the objective of \eqref{eqn-3}. Suppose that \me{t_k>Q_k} for a certain \me{k} at the optimum, i.e., \me{-{v}_k=t_k-Q_k>0}. Then there exists \me{\delta_k>0} such that \me{-{v}^{\prime}_k=t^{\prime}_k-Q_k<-{v}_k} where \me{t^{\prime}_k=t_k-\delta_k}. Since \me{\|\tilde{\mbf{v}}\|_q=\| |\tilde{\mbf{v}}| \|_q=\||-\tilde{\mbf{v}}|\|_q}, this means that the newly created vector \me{\mbf{t}^{\prime}} achieves a strictly smaller objective which contradicts with the fact that the optimal solution has been obtained. The choice of \me{\ell_q} norm used in the objective function \cite{berry2004cross,qps_cioffi} alters the priorities for the queue deviation function as follows.
\begin{itemize}
\item \me{\ell_{1}} norm results in greedy allocation, \textit{i.e}, emptying the queue of users with good channel states before considering the users with worse channel conditions. As a special case, it is easy to see that \eqref{eqn-3} reduces to the \ac{WSRM} problem when the queue size is large enough for all users.
\item \me{\ell_{2}} norm prioritizes users with a higher number of queued packets before considering the users with a smaller number of backlogged packets. For example, it would be more ideal for the delay limited scenario when the packet arrival rate of users are almost similar, since the number of queued packets waiting in the buffer is proportional to the transmission delay, by following Little's law \cite{neely2010stochastic}.
\item \(\ell_{\infty} \) norm minimizes the maximum number of queued packets among users with the current transmission, thereby providing queue fairness by allocating resources proportional to the number of backlogged packets.
\end{itemize}