### 2015-11-24 Response/Response 3/reviewer_response.tex
updated and finalized

\documentclass[10pt,letterpaper,onecolumn]{article}

\usepackage[left=1.0in,right=1.0in,top=1in,bottom=1in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage[dvipdfmx]{graphicx}
\usepackage{acronym,etoolbox,color,multicol,xcolor}
\usepackage{subeqnarray,multirow,cite,array,setspace,fixltx2e,lineno,verbatim,mathtools,pgfplots,subfig}

\input{./../../Report/shortcuts}
\input{./../../Report/acronyms}

\DeclareGraphicsExtensions{{eps}}
\graphicspath{{Figures/}}
\renewcommand \thesection{\Roman{section}}
\renewcommand \thesubsection{\arabic{section}.\arabic{subsection}}

\renewcommand{\thefigure}{\Roman{figure}}
\renewcommand{\thesubfigure}{\alph{subfigure}}
\renewcommand{\theequation}{\roman{equation}}

\begin{document}

	\title{Reviewers Comments \& Authors Replies}

	\date{}
	\maketitle

	\begin{tabular}{p{1.25in}p{4.25in}}
		\textbf{Manuscript No.} & Paper T-SP-18051-2014, submitted to \emph{``IEEE Transactions on Signal Processing"} \\ \\
		\textbf{Title} & ``Traffic Aware Resource Allocation Schemes for Multi-Cell MIMO-OFDM Systems" \\ \\
		\textbf{Authors} & Ganesh Venkatraman, Antti T\"{o}lli, Markku Juntti, and Le-Nam Tran
	\end{tabular}

	\vspace{0.35in}
	The authors are grateful to the associate editor for giving us an additional opportunity to revise and correct our convergence analysis. In this regard, we would like to give a special thank to Review 3 who pointed out the issues with our convergence analysis of the proposed solutions. In this revision, we have mainly focused on providing a more mathematically rigorous  proof of the convergence analysis of the proposed iterative algorithms.
	%\input{additions}
	Towards this end the following changes have been made to the revised manuscript.
	\begin{enumerate}
		\item We have restated the convergence results of the centralized algorithms in this revision, claiming that any accumulated point (i.e., a limit point) is a stationary point. The proof is essentially based on the convergence of a subsequence of iterates. We have rewritten Appendix A-D according to these changes. In particular, to discuss the convergence of the sequence of iterates generated by the centralized method, we use a unified superscript index \eqn{t} to refer to the step index of both \ac{AO} and \ac{SCA} procedures (\textit{i.e.}, index \eqn{t} denotes \ac{SCA} and \ac{AO} update step (\eqn{k,i})). Using this index \eqn{t}, we denote iterate \eqn{\ma^t} as a stacked vector of the minimizer obtained by solving convex subproblem \eqref{con-m} at each step \eqn{t} as \eqn{\mat{t} = \mbf{z}^{(i)}_k} for some \ac{SCA} and \ac{AO} iteration. Using this notation, we have revised our convergence discussions in Appendices A-C and A-D.		
		\item We have updated the convergence of distributed algorithms in a similar way.
	\end{enumerate}

	We have  also revised some parts of text to support the changes mentioned above and to further improve the readability. Specifically, the following parts of the revision have been rewritten.
	\begin{itemize}
		\item Page 6, paragraph after (28): We have elaborated the discussions on showing that the feasible set is bounded.
		\item Page 8, last paragraph in Section IV-B: We discussed the conditions under which the overall algorithm converges upon using the distributed schemes in each \ac{SCA} step.
		\item Last sentence after (51) on page 14: We emphasized the convergence of objective sequence only in each \ac{AO} step. 
		\item Last sentence after (60) on page 15: We have rewritten the convergence claim on the sequence of iterates generated by the Algorithm 2 similar to that of centralized method.
	\end{itemize}

	\vspace{1eM}
	In what follows, the comments are listed, each followed immediately by the corresponding reply from the authors. The reviewers questions in the revised manuscript are highlighted in blue color and the authors responses are presented in black. Unless otherwise stated, all the numbered items (figures, equations, references, citations, etc) in this response letter refer to the revised manuscript. All revisions in the manuscript are highlighted in blue color.
	\newpage
	\section*{Response to first reviewer comments}
	\input{response-1}

	\newpage
	\section*{Response to second reviewer comments}
	\input{response-2}

	\newpage
	\section*{Response to third reviewer comments}
	\input{response-3}

	\newpage
	\section*{Response to fourth reviewer comments}
	\input{response-4}

\end{document}