### 2016-02-08 IEEE TSP - Incl Bios/Report/conclusions.tex

In this paper, we have addressed the problem of allocating downlink space-frequency resources to the users in a multi-cell \ac{MIMO} \ac{IBC} system using \ac{OFDM}. The resource allocation is considered as a joint space-frequency precoder design problem since the allocation of a resource to a user is obtained by a non-zero precoding vector. We have proposed the \ac{JSFRA} scheme by relaxing the nonconvex constraint by a sequence of convex subsets using the \ac{SCA} for designing the precoders to minimize the total number of user queued packets. Additionally, an alternative \ac{MSE} reformulation approach has been proposed by using the \ac{SCA} to address the nonconvex constraints for a fixed \ac{MMSE} receivers. We also proposed various methods to decentralize the precoder designs for the \ac{JSFRA} problem using primal and \ac{ADMM} methods. Finally, we proposed a practical iterative algorithm to obtain the precoders in a decentralized manner by solving the \ac{KKT} expressions of the \ac{MSE} reformulated problem. The proposed iterative algorithm requires few iterations and limited signaling exchange between the coordinating \acp{BS} to obtain the efficient precoders for a given number of iterations. Numerical results are used to compare the performance of the proposed schemes.